---
title: "Data Management & Programming Project"
author: "Adrien BOUISSON & Ghina MEZHER"
format: html
warning: false
echo: false
message: false
editor: 
  markdown: 
    wrap: sentence
---

```{r}
#Calling all necessary libraries
#| warning: false
#| message: false 
here::i_am("project-BM.Rproj")
library(here)
library(dplyr)
library(ggplot2)
library(tidyr)
library(vroom)
library(knitr)
```
```{r message=FALSE}
# Attributing our databases to variables
#| message: false
dep_pub_raw<-vroom(here("dep-pub-raw.csv"),delim=",")
dep_pub<-vroom(here("dep-pub-raw.csv"),skip= 4)
chomage_jeune<-vroom(here("unemployment.csv"), delim = ",")
```

# Our research question

## I. The importance and history of research in education

It is undeniable education profoundly shapes societal progress. From allowing humans to make world changing discoveries to multiplying the efficiency and use of physical capital; education lets each individual make use of their production power. Education is also considered a fundamental right, as it is stipulated in the universal declaration of Human Rights 1948, and a key element of development as seen in the 2000 MDGs (MDG n°2) and the 2015 SDGs (SDG n°4).

Today, economic research in education is considered a greatly developed field, highlighting the importance of public spending in education and the necessity for targeted funding and strategic policies. We owe the establishment of the field to the theory of human capital by Schultz, Becker and Mincer established in the 50's . What started as a simple causality relationship between additional years of education and wage differences was frivolously applied to empirical studies, and econometric models. This allowed researchers to understand countries' particular attributes and situations but also if lack of education should be considered a demand or a supply problem.

Investing in human capital, here in its educational form, is considered fundamental to the reduction of inequalities. Take for example woman's fertility. Research has found that, in developing countries, one additional year of schooling reduces woman's fertility rate by 10%, enabling a much needed demographic transition. This causality relationship can be explained by the increase of revenue on the labor market through additional years of schooling, allowing women to take on better jobs and increasing the opportunity cost of having a child.

## II. The link between education and unemployment

The polarization of the labor market has become a clear and urgent issue. The dreaded phenomenon represents, for the middle class, the end of social mobility or as some might call it "The American/Capitalistic Dream". Polarization of the labor market defines the historical decline of the rate of jobs that lie in the middle of the distribution of qualification (France Stratégie 2020). The prosperity of machinery and innovation has led to the decrease of basic qualification demanding jobs in favor of highly qualified jobs or low qualified jobs. This phenomenon has revealed itself particularly challenging for under-qualified youth, whose unemployment rate has increased in the past few years.

In order to form the subject of our study we begun with the same idea as the theory of capital : how can altering the offer of education impacts individuals' status on the labor market.

However, in a will to understand the role the government must play when it comes to the offer in education we may ask ourselves : Should government leaders expect their educational spending to impact the unemployment rate ?

A higher unemployment rate could also prompt governments to allocate more funds to education as a way to respond to the need for training to boost employability, we therefore see a potential double causality between educational spending and unemployment. The potential reciprocal influence between educational investments and workforce participation unveils a complex interplay, emphasizing the need for nuanced analysis to comprehend the multifaceted impact of education on labor markets.

Ultimately, this exploration aims to inform policy decisions, offering insights into the interwoven pathways that shape both educational outcomes and employment prospects, fostering more holistic approaches towards societal development.

All these elements lead us to ask the following question: How can public spending on education help reduce youth unemployment in the world?

To answer our question, we will first look at the general trends in public spending and youth unemployment since 1970. We will then show that education spending is essential to reducing youth unemployment. Finally, we'll see that this is not enough, and that the reduction in the unemployment rate is made possible by other factors such as the country's development, rigidity on the labor market, etc.

# Data set description

Our topic of study in this project tackles the impact of education spending on the youth unemployment rate. For that, we have decided to take our databases from sites such as the World Bank and the International Labor Organization (ILO), as these are major global institutions known primarily for their availability and accessibility. For example, World Bank compiles and makes available a wide range of data on global economic development such as expenditures in school per country. 

Thanks to the richness of information in our panel ILO data base and panel World Bank database we will be able to study the impact of public educational spending on the educational aspect of human capital. With the detailed information from various countries and on a big date range, we are able to study the impact on a micro and macro level. The micro impact can be observed by studying each country on its own and observing the variations throughout the years. This also makes comparing countries possible; we could therefore observe the impact on developing countries and developed countries. We can also compare the effects of education spending on the youth unemployment rate within countries, i.e. between men and women, and between countries, i.e. the macroeconomic impact of education spending.

We would like to highlight the real and empirical impact of the augmentation of educational spending on level of education thanks to the "classif2" variable now called "Education". Through this potential correlation and causality we could understand the impact of a budgetary politic in education on unemployment. The underlying assumption is that a positive correlation between increased educational spending and higher education levels could translate into a subsequent decline in unemployment figures over the ensuing years. Therefore the project endeavors to decipher how these factors (public spending and education levels) interplay in the context of unemployment rates.

Our databases are dense, which requires delimiting our subject such as the period on which we are going to be studying or the variables that we consider to be most relevant.

## Description of the World bank database

First, many of the databases provided by the World Bank include time series data, allowing to analyze trends and changes over time. It could be essential for understanding the dynamics of variables. Then, the World Bank is known for maintaining high standards in databases. There is data reliability, World Bank data is generally subject to rigorous standards and quality controls. As part of our project, the database from the World Bank allows us to make international comparisons by collecting data from many countries, such as school expenditure for example.

Finally, by analyzing World Bank data on school expenditures, we can merge this database with the youth unemployment rate to study the correlation (or causality) of education spending on the unemployment rate.

It means that we could observe that the higher education spending led to the lower youth unemployment rate, or that the high youth unemployment rate can be explained by lower spending on education.

We have described below the necessary information of our database from the World Bank, which we have summarized in a table:

```{r echo:false}

#Our dep_pub_raw variable contains our dataset as it is wrongly read by R. We must keep it in order to retrieve information on the origin of the dataset and on its last update date as they are contained in the first 4 lines that make R reading our dataset complicated.
dep_pub|>
  summarise("Origin"="Table 1 : World bank","Last update"= dep_pub_raw[[1,2]],
            "Nb of countries observed"=n(),
            "Indicator"=dep_pub[[1,3]],
            "First date"=colnames(dep_pub)[5],
            "Last date"=colnames(dep_pub)[67],
            )|>
  knitr::kable()
```

### Variable description

-   "Country Name"/"Country Code": String value giving us an indicator for each country to classify values

-   "Indicator Name"/"Indicator code": String value giving us the indicator studied wich is the government expenditure on education as % of GDP

-   "Date": each column referring to a date for the % of GDP for government education spending going from 1960 to 2022

## Description of the International Labor Organisation database

Then, the International Labor Organization (ILO) provides a wealth of data related to labor and employment. The ILO sets international standards for labor and labor statistics. These standards guarantee a consistent methodology for the collection and presentation of data on the labor market, including the unemployment rate.This site therefore corresponds perfectly to our project.

ILO data offers comprehensive insights into the global labor marker, especially unemployment in the world, and youth unemployment (15-29 years old). Like the World Bank, the ILO collects and compiles data on youth unemployment from a wide range of countries, allowing for international comparisons which is perfect for our project. Both sites making international comparisons, it allows for an interesting world level analysis of our subject of study.

We have described below the necessary information of our database from the ILO, which we have summarized in a table:

### A table summary of our ILO database

```{r}
# Our database chomage_jeune contains a huge number of observations with 11 variables, so we'll remove the missing NA values to reduce the number of observations.

table_chomage<-colnames(chomage_jeune)
chomage_jeune|>
  summarise("Origin"="Table 2 : ILO",
            "Nb of variables"=ncol(chomage_jeune),
            "Nb of countries"= length(unique((chomage_jeune$ref_area))),
            "Sex groups"= length(unique((chomage_jeune$sex))),
            "Age groups "= length(unique((chomage_jeune$classif1))),
            "Number of dates" = length(unique((chomage_jeune$time))), 
            "First date"=min(chomage_jeune$time),
            "Last date"=max(chomage_jeune$time)
            )|>
  knitr::kable()
```

```{r}
#We rename the column to better understand the variables
colnames(chomage_jeune)[colnames(chomage_jeune) == "ref_areas"] <- "Country"
colnames(chomage_jeune)[colnames(chomage_jeune) == "classif1"] <- "Age groups"
colnames(chomage_jeune)[colnames(chomage_jeune) == "classif2"] <- "Education"
colnames(chomage_jeune)[colnames(chomage_jeune) == "obs_value"] <- "Youth unemployment (%)"
chomage_jeune <- chomage_jeune |> 
  mutate(`Age groups`=case_when(`Age groups`=="AGE_YTHBANDS_Y15-19" ~ "15-19",
                                `Age groups`=="AGE_YTHBANDS_Y15-29" ~ "15-29",
                                `Age groups`=="AGE_YTHBANDS_Y20-24" ~ "20-24",
                                `Age groups`=="AGE_YTHBANDS_Y25-29" ~ "25-29",
         TRUE ~`Age groups`))
chomage_jeune<-chomage_jeune|>
  mutate(`sex`=case_when(`sex`=="SEX_M"~"Male",
                       `sex`=="SEX_F"~"Female",
                       `sex`=="SEX_O"~"Other",
                       `sex`=="SEX_T"~"Total",
                       TRUE~`sex`))

chomage_jeune<-chomage_jeune|>
  mutate(`Education`=case_when(`Education`=="EDU_AGGREGATE_TOTAL"~"Total",
                       `Education`=="EDU_AGGREGATE_LTB"~"Less than basic",
                       `Education`=="EDU_AGGREGATE_BAS"~"Basic",
                       `Education`=="EDU_AGGREGATE_INT"~"Intermediate",
                       `Education`=="EDU_AGGREGATE_ADV"~"Advanced",
                       `Education`=="EDU_AGGREGATE_X"~"Level not stated",
                       TRUE~`sex`))
```

### Main variables description

"Ref area": String value giving us each country's code, for example "AFG" for Afghanistan. It can refer to countries, geographic regions, groups of countries (by income level or others) or the world

"sex": string value giving us the sex, three categories: total, male, female & other. The breakdown by sex and the items of this breakdown.

"classif1": string value giving us the youth range, three categories: 15-19, 20-24, 25-29, 15-29, for example "AGE_YTHBANDS_Y15-29","Age (Tranches des jeunes): 15-29"

"classif2": string value giving us the level of education of observed category, 6 categories are mentioned: Total (values for all categories), Less than basic, basic, intermediate, advanced and level not stated

"time": string value giving us the date of the observation

"obs_value": observed value according to each category

```{r}
# Cleaning the WB database
#In order to clean our world bank database we must delete columns/variables who have a unique value and dates that only have NA values.
#We rename our "Country Code" variable to "Country" and pivot our data frame in order to have a "time" variable to match the ILO database.

dep_pub_cleaning<-function(df){
  single_value<-function(my_var){
  (my_var|>unique()|>length())==1
  }
  country_codes<-df|>distinct(`Country Name`, `Country Code`)
  df_clean<-df|>
    select(-`Country Name`)|>
    rename(Country=`Country Code`)|>
    select(!where(\(.x)all(is.na(.x))))|>
    select(!where(single_value))|>
    pivot_longer(cols = -Country, names_to = "time", values_to = "Public Spending (% of GDP)")
}

dep_pub_clean<-dep_pub_cleaning(dep_pub)
```

```{r}
# Cleaning the ILO database
#We remove some column which will not used for the projet like "Source", "Indicator", "obs_status", "note_indicator", "note_source", and we create a new database called chomage_jeune_clean.These are just codes used by the ILO to classify data.
chomage_jeune_cleaning<-function(df){
  single_value<-function(my_var){
  (my_var|>unique()|>length())==1
  }
  df_clean<-df|>
    filter(sex != "Other")|>
    rename(Country=ref_area)|>
    select(!where(\(.x)all(is.na(.x))))|>
    select(!where(single_value)) |>
    select(-"source", -"obs_status", -"note_source", -"note_indicator")
}
chomage_jeune_clean<-chomage_jeune_cleaning(chomage_jeune)
```

```{r}
# Merging our databases
# We have a new database merging by both database. In the following we will maily use it. 
dep_chomage <- merge(dep_pub_clean, chomage_jeune_clean, by = c("Country", "time"), all = FALSE)
dep_chomage$`Public Spending (% of GDP)`<-as.numeric(dep_chomage$`Public Spending (% of GDP)`)
dep_chomage$time<-as.numeric(dep_chomage$time)
```

## A table summary of our database

We've merged our 2 databases in order to match them by country (variable "Country") and date (variable "time").

```{r}
dep_chomage|>
  summarise("Title"="Table 3 : Education spending and youth unemployment rate by country over time",
            "Nb of observations"=n(),
            "Nb of variables"=ncol(dep_chomage),
            "Nb of countries"=length(unique(dep_chomage$Country)),
            "Age groups"=length(unique(dep_chomage$`Age groups`,na.rm=TRUE)),
            "Time frame"=paste(min(dep_chomage$time),"-",max(dep_chomage$time)))|>
  knitr::kable()
```

# Data Analysis

## I. Public spending and youth unemployment in the world

### 1) Public spending on education since 2014 in the world 

To begin our analysis, we'll be looking at how public spending on education has evolved around the world, using the ggplot function to introduce graphs. This will enable us to look at general trends, and to see which countries have the highest and lowest educational expenditure.

```{r}
mean_by_date <- dep_chomage |>
  group_by(time) |>
  summarise(mean_ps= mean(`Public Spending (% of GDP)`, na.rm = TRUE))
  
ggplot(mean_by_date,aes(x=time,y =mean_ps)) +
  geom_col() +
  ylim(0,7)+
  labs(x = "Date", y = "World mean Public Spending (% of GDP)", title = "Figure 1 : Mean Public Spending Over Time", caption = "Source: World Bank")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Therefore we can see public spending varies through time with a particular low in 1996 and 1997. Over the past few years, the average public spending on education has hovered around 4.8% of GDP with a decline starting 2021 we might imagine a reorientation of public spending to other sectors like health, security and energy following the pandemic and the war in Ukraine. It is therefore interesting to conduct a more in depth examination of public expenditure trends since 2014, in order to define which countries contribute to elevating this average and those that contribute to lowering it.

```{r}
dep_pub_14_22<-dep_pub_clean|> filter(time>=2014& time<=2022)
  
ggplot(dep_pub_14_22, aes(x=time,y=`Public Spending (% of GDP)`))+
  geom_line(color="black") +  
  geom_point(color="red") +
  geom_text(data = dep_pub_14_22 |> filter(`Public Spending (% of GDP)` > 9),
           aes(label = Country), hjust = 0, vjust = 0, size = 3, color = "black")+
  labs(title = "Figure 2 : Public spending on education trends (2014-2022)",
       x = "Time",
       y = "Public spending (% of GDP)", 
       caption = "Source: World Bank")
```
Thanks to this graph, we have the evolution of public spending on education over the period 2014-2022. The dots represent the countries, we can therefore see that the average public expenditure in 2021 is elevated by Kiribati and the Marshall Islands' public spending on education. These two countries have also relatively high education spending in this time span, also adding to that list Namibia and the Federated States of Micronesia. Therefore examining the influence of public spending on the youth unemployment rate in countries  situated in Africa (for the case of Namibia) would be a compelling study taking into account that the MDGs and SDGs on education focused on development projects in the region.

### a) The largest share of GDP for public spending in education 
```{r}
country_highest_spending <- dep_chomage |>
  group_by(time) |>
  slice(which.max(`Public Spending (% of GDP)`)) |>
  ungroup()

country_highest_spending_filter <- country_highest_spending |>
  filter(time >= 2000 & time <= 2022)

ggplot(country_highest_spending_filter, aes(x = time, y = `Public Spending (% of GDP)`, fill = Country)) +
  geom_bar(stat = "identity") +
  labs(title = "Figure 3 :Countries with Highest Public Spending (% of GDP) by Year",
       x = "Year",
       y = "Public Spending (% of GDP)",
       fill = "Country", 
       caption = "Source: World Bank") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Countries that devote a large share of their GDP to education spending are : Samoa américaines (ASM), Cuba (CUB), Micronésie, Etats fédérés (FSM), Kiribati (KIR), Lesotho (LS0), Marshall's Islands (MHL), Namibie (NAM). 

### b) The smallest share of GDP for public spending 

```{r}
country_smallest_spending <- dep_chomage |>
  group_by(time) |>
  slice(which.min(`Public Spending (% of GDP)`)) |>
  ungroup()

country_smallest_spending_filter <- country_smallest_spending |>
  filter(time >= 2000 & time <= 2022)

ggplot(country_smallest_spending_filter, aes(x = time, y = `Public Spending (% of GDP)`, fill = Country)) +
  geom_bar(stat = "identity") +
  labs(title = "Figure 4 :Countries with smallest Public Spending (% of GDP) by Year",
       x = "Year",  y = "Public Spending (% of GDP)",
       fill = "Country", 
       caption = "Source: World Bank") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Countries that devote a large share of their GDP to education spending are : Africa Eastern and Southern (AFE), Central African Republic (CAF), Zambia (ZMB), South Sudan (SSD), Sudan (SDN), Gambia (GMB), etc.. The majority of countries that spend the least on education are in Africa. 

### 2) Youth unemployment in the world 

In this section, we'll take a look at the evolution of the global youth unemployment rate over the period 2014-2022. Taking the same period as public spending will allow us to account for a potential correlation (or even causality) between public spending and the unemployment rate in a second part. We therefore proceed in the same way as above, taking the most developed countries on the one hand, and developing countries on the other. Here, we analyze only the major trends.

Employment is a means of accessing the wealth produced through work, but not all work is employment. It is a source of income. While work designates the activity.

Unemployment refers to the deprivation of a job. An unemployed person is not deprived of work, but of employment. For the person who holds it, employment is a source of income through participation in the economic distribution of the wealth produced.

According to ILO, an unemployed person is a person aged 15 or over who simultaneously meets three conditions: being unemployed during a given week; being available to take up a job within two weeks; having actively looked for a job in the last four weeks or having found one starting in less than three months.

In the wake of worsening youth unemployment in Europe, a debate has recently emerged about the instrument used at the ILO: should rates or ratios be used to express youth unemployment?

The difference lies in the denominator: only the young workforce - those working or looking for work - is taken into account in the case of the rate, as opposed to the entire 15-24 age group - including full-time students - in the case of the ratio.

So, if we look at 200 young people, 100 of whom are students and 50 employed, the unemployment rate is 50 percent, while the ratio is 25 percent.

Applying a ratio can be useful for comparing levels of youth unemployment between countries, because there are significant differences in the way countries count youth participation in the workforce. In the case of ILO database we have a unemployment rate which which already provides an overview of our research project. 

```{r}
mean_by_date_unemployment <- dep_chomage %>%
  group_by(time) %>%
  summarise(mean_unemployment = mean(`Youth unemployment (%)`, na.rm = TRUE))

ggplot(mean_by_date_unemployment, aes(x = time, y = mean_unemployment)) +
  geom_col() +
  labs(x = "Date", y = "World mean Youth Unemployment (%)", title = "Figure 5 : Mean Youth Unemployment Over Time", caption = "Source: ILO") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
We can see that the global youth unemployment rate has been rising for several decades. At the same time, we have seen a decrease in the number of young people unemployed.  
```{r}
dep_chom_14_22 <- chomage_jeune_clean %>% 
  filter(time >= 2014 & time <= 2022, `sex` == "Total")

ggplot(dep_chom_14_22, aes(x = time, y = `Youth unemployment (%)`)) +
  geom_point(color = "red") +
  geom_text(data = dep_chom_14_22 %>% filter(`Youth unemployment (%)` > 75),
            aes(label = Country), hjust = 0, vjust = 0, size = 3, color = "black") +
  labs(title = "Figure 6 :Youth Unemployment Trends (2014-2022)",
       x = "Time",
       y = "Youth Unemployment (%)", 
       caption = "Source: ILO")
```
Through this graph we can see countries with highest youth unemployment rate at the top of the line. They are Kosovo, South Africa, Uruguay,Tuvalu and Greece. They do not appear as countries who spend the lowest % of GDP on education, it is therefore interesting to conduct a study on African countries in order to understand the potential causality or non-causality between public expenditure and youth unemployment.

### a) The highest youth unemployment rate - 15-19 & 20-24 years old 

Having seen which countries devote the most or least of their GDP to education spending, we're now going to look at which countries have the highest and lowest unemployment rates among young people, in this case 15-24 year-old. 

```{r}
youth_unemployment_highest <- dep_chomage %>%
  group_by(time) %>%
  filter(time >= 2000 & time <= 2022, `Age groups` %in% c("15-19", "20-24"), `Education`== "Total", sex == "Total") %>%
  slice(which.max(`Youth unemployment (%)`)) %>%
  ungroup()

ggplot(youth_unemployment_highest, aes(x = time, y = `Youth unemployment (%)`, fill = Country)) +
  geom_bar(stat = "identity") +
  labs(title = "Figure 7 : Countries with Highest Youth Unemployment by Year",
       x = "Year",
       y = "Youth Unemployment (%)",
       fill = "Country", 
       caption = "Source: ILO") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Countries with the highest youth unemployment rates are : Angola (AGO), Bosnia and Herzegovina (BIH), Djibouti (DJI), Spain (ESP), Saint Lucia (LCA), North Macedonia (MKD), South Africa (ZAF). 

### b) The lowest youth unemployment rate - 15-19 & 20-24 years old 

```{r}
youth_unemployment_lowest <- dep_chomage %>%
  group_by(time) %>%
  filter(time >= 2000 & time <= 2022, `Age groups` %in% c("15-19", "20-24"), `Education`== "Total", sex == "Total") %>%
  slice(which.min(`Youth unemployment (%)`)) %>%ungroup()

ggplot(youth_unemployment_lowest, aes(x = time, y = `Youth unemployment (%)`, fill = Country)) +
  geom_bar(stat = "identity") +
  labs(title = "Figure 8: Countries with Lowest Youth Unemployment by Year",
       x = "Year",
       y = "Youth Unemployment (%)",
       fill = "Country",
       caption = "Source: ILO") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")
```
Countries with the lowest youth unemployment rates are : Switzerland (CHE), Guatemala (GTM), Cambodia (KHM), Liberia (LBR), Moldova (MDA), Madagascar (MDG), Mali (MLI), Malawi (MWI), Niger (NER), Pakistan (PAK), Peru (PER), Rwanda (RWA). 

We can see from this graph that the place where youth unemployment is highest is predominantly in Africa. Public spending can therefore have an impact on the youth unemployment rate. But we can also see that some African countries have the lowest unemployment rates, so education spending isn't the only explanatory variable. 

### 3) General observation of public spending in education and youth unemployment 

In order to have a more in depth analysis of general trends and understand the different impact of public expenditure on sexes, age groups and level of education we must select one  year to study out of our panel data. The code becomes quite simple to alter in order to study another date by replacing the year 2021 that we have chosen with whichever year is of interest to the researcher. 

We have chosen the year 2021 as it is the year where public expenditure started to recede. This allows us to observe a potential resurge of youth unemployment. Add onto that the high public spending of some countries that year, we will be able to observe their impact on youth unemployment according to sex and level of education.

#### a) general observations
```{r}
dep_chomage |> filter(time==2021) |>
  ggplot(aes(x=`Public Spending (% of GDP)`,y=`Youth unemployment (%)`)) +
  geom_hex()+
  xlim(0,17)+
  facet_grid(~sex)+
  geom_rect(aes(xmin=1.5,xmax=3.5,ymin=27,ymax=75),color="red",fill=NA)

dep_chomage$time<-as.numeric(dep_chomage$time)
dep_chomage |> filter(time==2021) |>
  ggplot(aes(y=`Public Spending (% of GDP)`,x=`Youth unemployment (%)`)) +
  geom_hex()+
  ylim(0,17)+
  facet_grid(~Education)+
  geom_rect(aes(ymin=1.5,ymax=8,xmin=27,xmax=77),color="red",fill=NA)
```
We can therefore see a small disparity in the distribution of points between the subplots of each of these two graphs. When it comes to the genders, women tend to have a higher youth unemployment rate than men for the same public spending in % of GDP. When it comes to level of education, for the same amount of public spending in education, the youth unemployment rate tends to be higher for a level of education stated as "Basic" than "Advanced". We would like to conduct a more in depth study of the impact of public expenditure on the level of education as higher education levels seem to better benefit from public spending. 

Countries with high public spending tend to have overall lower youth unemployment rate might it be for the different sexes or level of education. The disparity in youth unemployment between men and women and different levels of education is less prominent for countries with higher public spending. 

The graphs remain quite hard to read as each point represents one country for one age range, it is therefore interesting to take on the analysis of one representative country for each of our later sections to better see how public spending in education impacts men and women differently. 

#### b) impact of public spending on education level
```{r}
dep_chomage|> filter(time==2021)|>
  ggplot(aes(x=`Public Spending (% of GDP)`,y=`Youth unemployment (%)`,color=Education)) +
  geom_point(alpha=0.5)+
  facet_grid(~`Age groups`)+
  geom_vline(xintercept = 7.8, col = "red",alpha=0.4)
```
This general observation allows us to take into account the impact of public spending on education by Age groups and by level of education, for the year 2021. 

In order to read this graph we could take a look at a specific value of PUblic Spending, for example 7.8% of GDP. This is considered a high public spending in education. We can therefore see which levels of education are benefiting from this public expenditure. We can see that for such a public spending and for the different age ranges, advanced level of education tends to have a lower youth unemployment rate. For lower public spending we are startled to see an "Advanced" red dot with a 100% youth unemployment rate, this could be a specific case of a developing country where unemployment is hardly defined. 

Public spending therefore impacts individuals differently and according to their level of education. An individual that has an intermediate level at 15-19 and gets on the labor market will not benefit from the public expenditure. This is why for countries with higher publix spending we can see the points tend to stay lower as public expenditure promotes and helps individul stay in the school system. 


```{r}
mean_public<- mean(dep_chomage$`Public Spending (% of GDP)`, na.rm = TRUE)
dep_chomage$high_low <-ifelse(dep_chomage$`Public Spending (% of GDP)`> mean_public, "High", "Low")

ggplot(dep_chomage, aes(`Youth unemployment (%)`, fill = high_low)) +
  geom_histogram(binwidth = 2)+
  scale_fill_manual(values = c("darkgreen", "red"), name = "High/Low")
```
Here we divided our observations into 2 sections thanks to the creation of a High/Low dummy variable. Countries with "High" public spending in education have a % of GDP dedicated to education superior to the mean of our database. Here we can see that as youth unemployment rises, the count of countries with a high public spending decreases becoming smaller than the red bars that represent "Low" spending in education. 

## II. Public spending and youth unemployment in developed countries

Counter-intuitively, developed countries do not allocate the highest % of GDP to public spending on education. 

We'll be looking at public spending trends in developed countries such as the United States, France, Germany, Japan, Canada, the United Kingdom, Russia and Italy.

### 1) PUblic spending in education in developed countries: G8 countries 

```{r}
dep_chomage$time<-as.numeric(dep_chomage$time)
  
dep_chomage_g8 <- dep_chomage |>
  filter(!is.na(`Public Spending (% of GDP)`))|>
  filter(Country %in% c("USA", "FRA", "GER", "JAP", "CAN", "GBR", "RUS", "ITA"))

mean_g8 <- dep_chomage_g8 |>
  group_by(time) |>
  summarise(mean_public_spending = mean(`Public Spending (% of GDP)`, na.rm = FALSE))

ggplot(mean_g8,aes(x=time,y=mean_public_spending))+
  geom_point(na.rm = FALSE) +
   geom_line(na.rm=FALSE)+
    labs(title = "Figure 9 :Public spending on education trends for developped countries",
       x = "Time",
       y = "Mean public spending (% of GDP)", 
       caption = "Source: World Bank") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Since the 1970s, there has been a general decline trend in education spending in the G8 countries. This decline can be explained by the aging of the population, leading to a reallocation of public spending to health. Another factor may be government priorities.
The sudden increase in 1993 can be explained by the fall of the Berlin Wall and the end of the Cold War, which reduced trade between countries and hence GDP. As education spending remained unchanged, its share of GDP increased. In 2020, the decline in the share of education spending fell sharply as a result of the health crisis, which reduced both GDP and education spending. 

Moreover, this graph plots the average for G8 countries without taking into account the disparities between them, as we shall see in the following graph. 

```{r}
dep_chomage_g8_filtered <- dep_chomage_g8 |>
  filter(!is.na(`Public Spending (% of GDP)`)) |>
  filter(time >= 2014 & time <= 2022)
ggplot(dep_chomage_g8_filtered, aes(x = time, y = `Public Spending (% of GDP)`, color = Country, group = Country, shape = Country)) +
  geom_line(na.rm = FALSE) +  
  geom_point(na.rm = FALSE) + 
  labs(title = "Figure 10 :Public spending on education trends for G8 countries (2014-2022)",
       x = "Year",
       y = "Public spending (% of GDP)",
       color = "Country", 
       caption = "Source: World Bank") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Focusing solely on the G8 countries, we have assigned a color to each country, showing that the United States (pink), France (yellow) and Great Britain (green) are the countries that spend the most on education. We can see that USA's public spending remains around 5,5% of the GDP compared to Italy with 4,1 of GDP. So, if we take the general trend into account, we can see that for the selected countries we have a drop in public spending on education. Moreover, we don't have any information about public spending on education in Russia for 2021 and 2022. 

### 2) Unemployement rate in the developped countries : G8 countries

```{r}
dep_chomage$time<-as.numeric(dep_chomage$time)
  
dep_chomage_g8 <- dep_chomage |>
  filter(!is.na(`Youth unemployment (%)`))|>
  filter(Country %in% c("USA", "FRA", "GER", "JAP", "CAN", "GBR", "RUS", "ITA"))

mean_g8_unemployment <- dep_chomage_g8 |>
  group_by(time) |>
  summarise(mean_youth_unemployment = mean(`Youth unemployment (%)`, na.rm = FALSE))

ggplot(mean_g8_unemployment,aes(x=time,y=mean_youth_unemployment))+
  geom_point(na.rm = FALSE) +
   geom_line(na.rm=FALSE)+
    labs(title = "Figure 11 : Youth unemployment in developped countries (G8)",
       x = "Year",
       y = "Mean youth unemployment rate (%)", 
       caption = "Source: ILO") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
dep_chomage_g8_filtered <- dep_chomage_g8 |>
  filter(!is.na(`Youth unemployment (%)`)) |>
  filter(time >= 2014 & time <= 2022)

ggplot(dep_chomage_g8_filtered, aes(x = time, y = `Youth unemployment (%)`, color = Country, group = Country, shape = Country)) +
  facet_wrap(~Country) +
  geom_line() +
  geom_point() +
  labs(title = "Figure 12 : Youth Unemployment Rate Trends for G8 Countries (2014-2022)",
       x = "Year",
       y = "Youth Unemployment (%)",
       color = "Country",
       caption = "Source: ILO") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
dep_chomage_g8_unemployment <- dep_chomage_g8 %>%
  filter(time >= 2014 & time <= 2022, sex == "Total")

ggplot(dep_chomage_g8_unemployment, aes(x = time, y = `Youth unemployment (%)`, color = Country, group = Country, shape = Country)) +
  geom_smooth(na.rm = FALSE) +  
  geom_point(na.rm = FALSE) + 
  labs(title = "Figure 12 Bis : Youth unemployment Rate trends for G8 countries (2014-2022)",
       x = "Time",
       y = "Youth unemployment (%)",
       color = "Country", 
       caption = "Source: ILO") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
This pair of graphs allows us to compare the trends in youth unemployment rates among G8 countries. The United States has the lowest youth unemployment rate.

### 3) Youth unemployment according to Public spending  for G8 countries 

#### a) general observations according to education level and sex 
```{r}
dep_chomage_g8 |> filter(time==2021) |>
  ggplot(aes(x=`Public Spending (% of GDP)`,y=`Youth unemployment (%)`)) +
  geom_hex()+
  xlim(0,7)+
  facet_grid(~sex)+
  geom_rect(aes(xmin=3,xmax=6,ymin=27,ymax=75),color="red",fill=NA)

dep_chomage_g8$time<-as.numeric(dep_chomage_g8$time)
dep_chomage |> filter(time==2021) |>
  ggplot(aes(y=`Public Spending (% of GDP)`,x=`Youth unemployment (%)`)) +
  geom_hex()+
  ylim(0,17)+
  facet_grid(~Education)+
  geom_rect(aes(ymin=1.5,ymax=8,xmin=27,xmax=77),color="red",fill=NA)
```
When we examine public spending and the youth unemployment rate by gender for G8 countries, we can observe that, for the G8 countries, any increase in public spending results in a reduction in the unemployment rate, although this result is less evident for women.

#### b) according to education and age range 
```{r}
dep_chomage_g8_filtered|>
  ggplot(aes(x=`Public Spending (% of GDP)`,y=`Youth unemployment (%)`,color=Education)) +
  geom_point(alpha=0.5)+
  facet_grid(~`Age groups`)+
  geom_vline(xintercept = 6.1, col = "red",alpha=0.4)
```
For the 15-19 age group, we can see that the increase in public spending allocated to education has a more significant effect for a higher level of education.

### III - Public spending in developping countries - on the African continent

We'll be looking at public spending trends in Africa such as : Africa Eastern and Southern (AFE), Burundi (BDI), Burkina Faso (BFA), Belize (BLZ), Central African Republic (CAF), Congo - Democratic Republic of Congo (COD), etc..

#### a) General trends in public spending on education (% of GDP)

```{r}
dep_chomage_afr <- dep_chomage %>%
  filter(Country %in% c("AFE", "BDI", "BFA", "BLZ", "CAF", "COD", "MDG", "MLI", "NER", "MWI", "RWA", "SSD", "ZMB", "DJI", "COD", "LBR", "SDN", "SEN", "DAZ", "BWA", "TCD", "EGY", "ERI", "ETH", "GMB", "KEN", "LBY", "MAR", "NAM", "NGA", "SLE", "ZAF", "TGO", "TUN", "YEM", "ZWE"))

dep_chomage_africa_filtered <- dep_chomage_afr |>
  filter(!is.na(`Public Spending (% of GDP)`)) |>
  filter(time >= 2014 & time <= 2022)
ggplot(dep_chomage_africa_filtered, aes(x = time, y = `Public Spending (% of GDP)`, color = Country, group = Country, shape = Country)) +
  geom_line(na.rm = FALSE) +  
  geom_point(na.rm = FALSE) + 
  labs(title = "Figure 13 : Public spending on education trends for African countries (2014-2022)",
       x = "Year",
       y = "Public spending (% of GDP)",
       color = "Country", 
       caption = "Source: World Bank") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")
```
In comparison with the graph showing the evolution of the share of public expenditure in GDP for the G8 countries (Figure 10), we could see disparities between countries but this disparity is even more pronounced for African countries going up to 'at 'nearly 8 percentage points difference (Namibia vs Madagascar). 

### 2) General trends in unemployement rate

```{r}
# We keep African countries which have complete data 

dep_chomage_afr_unemployment <- dep_chomage |>
  filter(Country %in% c("BDI", "BFA", "BLZ", "COD", "MDG", "MLI", "NER", "MWI", "RWA", "ZMB", "COD", "LBR", "SEN", "BWA", "EGY", "ETH", "GMB", "KEN", "LBR", "NAM", "NGA", "SLE", "ZAF", "TGO", "TUN", "YEM", "ZWE")) |>
  filter(!is.na(`Youth unemployment (%)`)) |>
  filter(time >= 2014 & time <= 2022)

mean_afr_unemployment <- dep_chomage_afr_unemployment |>
  group_by(time) |>
  summarise(mean_youth_unemployment_afr = mean(`Youth unemployment (%)`, na.rm = FALSE))

ggplot(mean_afr_unemployment,aes(x=time,y=mean_youth_unemployment_afr))+
  geom_point(na.rm = FALSE) +
   geom_line(na.rm=FALSE)+
    labs(title = "Figure 14 : Youth unemployment in African countries (2014-2022)",
       x = "Year",
       y = "Mean youth unemployment rate (%)", 
       caption = "Source: ILO") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Youth unemployment in Africa is higher than in the rest of the world. But some African countries have a higher unemployment rate, so we will see the evolution of the youth unemployment rate by separating the countries. 
```{r}
dep_chomage_afr_unemployment_1 <- dep_chomage |>
  filter(Country %in% c("BLZ", "MDG", "MLI", "MWI", "RWA", "LBR", "SEN", "EGY", "GMB", "KEN", "LBR", "NAM", "NGA", "SLE", "ZAF", "TGO", "TUN", "YEM", "ZWE")) |>
  filter(!is.na(`Youth unemployment (%)`)) |>
  filter(time >= 2014 & time <= 2022)

ggplot(dep_chomage_afr_unemployment_1, aes(x = time, y = `Youth unemployment (%)`, color = Country, group = Country, shape = Country)) +
  geom_smooth(na.rm = FALSE) +  
  geom_point(na.rm = FALSE) + 
  labs(title = "Figure 15 : Youth unemployment Rate trends for African countries (2014-2022)",
       x = "Time",
       y = "Youth unemployment (%)",
       color = "Country", 
       caption = "Source: ILO") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")
```
```{r}
dep_chomage_afr_unemployment_1 <- dep_chomage |>
  filter(Country %in% c("BLZ", "MLI", "MWI", "RWA", "EGY", "ZAF")) |>
  filter(!is.na(`Youth unemployment (%)`)) |>
  filter(time >= 2014 & time <= 2022)

ggplot(dep_chomage_afr_unemployment_1, aes(x = time, y = `Youth unemployment (%)`, color = Country, group = Country, shape = Country)) +
  facet_wrap(~Country) +
  geom_line() +
  geom_point() +
  labs(title = "Figure 15 BIS : Youth Unemployment Rate Trends for African Countries (2014-2022)",
       x = "Year",
       y = "Youth Unemployment (%)",
       color = "Country",
       caption = "Source: ILO") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")
```

### 3) Impact of public spending on education level on African countries

#### a) general observations

```{r}
dep_chomage_afr_unemployment_1 |> filter(time==2021) |>
  ggplot(aes(x=`Public Spending (% of GDP)`,y=`Youth unemployment (%)`)) +
  geom_hex()+
  xlim(0,17)+
  facet_grid(~sex)+
  geom_rect(aes(xmin=1.5,xmax=3.5,ymin=27,ymax=75),color="red",fill=NA)

dep_chomage_afr_unemployment_1$time<-as.numeric(dep_chomage_afr_unemployment_1$time)
dep_chomage |> filter(time==2021) |>
  ggplot(aes(y=`Public Spending (% of GDP)`,x=`Youth unemployment (%)`)) +
  geom_hex()+
  ylim(0,17)+
  facet_grid(~Education)+
  geom_rect(aes(ymin=1.5,ymax=8,xmin=27,xmax=77),color="red",fill=NA)
```

#### b) according to education
```{r}
dep_chomage_africa_filtered|>
  ggplot(aes(x=`Public Spending (% of GDP)`,y=`Youth unemployment (%)`,color=Education)) +
  geom_point(alpha=0.5)+
  facet_grid(~`Age groups`)+
  geom_vline(xintercept = 3.8, col = "red",alpha=0.4)
```

# Statistical test, linear regression, instrumental variables

The second part being a more descriptive part highlighting the major trends in the world and in specific geographical areas, such as Africa in our case. In this last part of our project we will look carefully at the different links that education spending and the youth unemployment rate have between them. For this we will carry out statistical tests such as the chi-square law, linear regressions.

Instrumental variables are used to solve the endogeneity problem in econometric models. We want to run a linear regression looking at the impact of education spending on the youth unemployment rate. In our project, we will use education expenditure as the explanatory variable and the youth unemployment rate as the explained variable.

To deepen our analysis, we will use instrumental variables, which are variables used to solve the endogeneity problem in econometric models. Endogeneity can arise when the explanatory variable is correlated with model errors, which can bias estimation results. Instrumental variables are therefore used to isolate the uncorrelated variation of the explanatory variable and use it as a proxy for that variable.

For the purposes of this project, we can therefore use the following instrumental variables: total government expenditure, number of schools per capita and population growth rate. This allows us to see whether these variables are not correlated directly with the youth unemployment rate, but are correlated directly with education spending.

## I) Correlation test for the world 

To understand the relationship between public spending on education and the youth unemployment rate, we will perform a correlation analysis.

### 1) Correlation

```{r}
correlation_result <- cor.test(dep_chomage$`Public Spending (% of GDP)`, dep_chomage$`Youth unemployment (%)`)
print(correlation_result) 
```
### 2) Linear Regression 

```{r}
linear_model <- lm(`Youth unemployment (%)` ~ `Public Spending (% of GDP)`, data = dep_chomage)
summary(linear_model)
```
#### Regression Model

The simple linear regression model can be represented as:

\[
\ Y_{t} = ß_{0} + ß_{1}X_{t} +  \varepsilon_{t}
\]

where:
- \(Y_{t}\): Youth unemployment rate at time \(t\).
- \(\beta_{0}\): Intercept term.
- \(\beta_{1}\): Coefficient for the variable \(X_{t}\).
- \(X_{t}\): Public spending variable at time \(t\).
- \(\epsilon_{t}\): Error term.

```{r}
## Graph representation 

scatter_plot_world <- ggplot(dep_chomage, aes(x = `Public Spending (% of GDP)`, y = `Youth unemployment (%)`)) +
  geom_point() +
  labs(
    title = "Figure 16 : Youth Unemployment vs Public Spending on Education",
    x = "Public Spending (% of GDP)",
    y = "Youth Unemployment (%)", 
    caption = "Source : ILO/World Bank"
  )

regression_line_plot <- scatter_plot_world +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  theme_minimal()

print(regression_line_plot)

```
When we take the global data, the regression line is almost straight, or slightly decreasing, this may suggest that for every increase (decrease) in education spending, the youth unemployment rate decreases (or increases) by constant manner. Furthermore, this may also suggest an absence of causal link between the two variables.

### 3) Chi deux test 

```{r}
chi_square_result <- chisq.test(table(dep_chomage$`Youth unemployment (%)`, dep_chomage$`sex`))
print(chi_square_result) 
```
The p-value of 0.03145 is less than the common significance level of 0.05. With a p-value below 0.05, we must reject the null hypothesis. In the context of a chi-square test for independence, this result suggests that there is evidence of an association between the variables Youth unemployment (%) and gender. The association is considered statistically significant.

## II) Statistical test - G8 countries 

### 1) Correlation test

```{r}
correlation_g8 <- cor.test(dep_chomage_g8_filtered$`Public Spending (% of GDP)`, dep_chomage_g8_filtered$`Youth unemployment (%)`)
print(correlation_g8)
``` 
### 2) Linear regression - G8 countries

```{r}
linear_model <- lm(`Youth unemployment (%)` ~ `Public Spending (% of GDP)`, data = dep_chomage_g8)
summary(linear_model)
```

```{r}
correlation_g8 <- cor.test(dep_chomage_g8_filtered$`Public Spending (% of GDP)`, dep_chomage_g8_filtered$`Youth unemployment (%)`)

matrix_g8 <- matrix(c(correlation_g8$estimate, correlation_g8$p.value), nrow = 1, dimnames = list(NULL, c("Correlation Coefficient", "P-Value")))

print(matrix_g8)
```

The correlation test provides valuable insights into the strength and direction of the linear relationship between public spending on education and the youth unemployment rate. The correlation coefficient and p-value help us assess the statistical significance of the observed correlation. In this case, with a negative correlation coefficient between public spending (% of GDP) and youth unemployment (%), this suggests that there could be a trend where higher levels of public spending are associated with lower rates of youth unemployment, and vice versa. 

```{r}
ggplot(dep_chomage_g8_filtered, aes(x = `Public Spending (% of GDP)`, y = `Youth unemployment (%)`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Figure 17 : Youth Unemployment vs Public Spending on Education in G8 countries",
       caption = "Source : ILO/World Bank")
```
The scatter plot with a fitted line visually represents the correlation between public spending and the youth unemployment rate. The blue line indicates the linear regression fit.

The correlation test and visualization help us understand if there is a significant correlation between public spending on education and the youth unemployment rate. The correlation coefficient close to 1 or -1 indicates a strong relationship, while a p-value less than 0.05 suggests statistical significance.

## III) Statistical test - African countries 

### 1) Correlation test 

```{r}
correlation_afr <- cor.test(dep_chomage_africa_filtered$`Public Spending (% of GDP)`, dep_chomage_africa_filtered$`Youth unemployment (%)`)
print(correlation_g8)
```
### 2) Linear regression - African countries 

```{r}
linear_model <- lm(`Youth unemployment (%)` ~ `Public Spending (% of GDP)`, data = dep_chomage_afr_unemployment)
summary(linear_model)
```
```{r}
correlation_afr <- cor.test(dep_chomage_africa_filtered$`Public Spending (% of GDP)`, dep_chomage_africa_filtered$`Youth unemployment (%)`)

matrix_afr <- matrix(c(correlation_g8$estimate, correlation_g8$p.value), nrow = 1, dimnames = list(NULL, c("Correlation Coefficient", "P-Value")))

print(matrix_afr)
```

```{r}
ggplot(dep_chomage_africa_filtered, aes(x = `Public Spending (% of GDP)`, y = `Youth unemployment (%)`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Figure Figure 18 : Youth Unemployment vs Public Spending in African countries",
       caption = "Source : ILO/World Bank")
```
When we take all African countries together, we can see a different correlation to the one previously calculated. In fact, it's positive, suggesting that the higher the public spending, the higher the youth unemployment rate. 

A RAJOUTER DESCRIPTION RAPIDE

## IV) Add instrumentales variables : stability of countries for example 

To further enhance the analysis, we can add additional explanatory variables such as gender and education level to our regression models. We will only do this for the G8 countries.

### 1) Linear regression with education and sex for G8 countries 

#### a) Education

```{r}
linear_model_g8 <- lm(`Youth unemployment (%)` ~ `Public Spending (% of GDP)` + Education, data = dep_chomage_g8_filtered)

summary(linear_model_g8)

scatter_plot_g8 <- ggplot(dep_chomage_g8_filtered, aes(x = `Public Spending (% of GDP)`, y = `Youth unemployment (%)`, color = Education)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(
    title = "Figure 18 : Youth Unemployment vs Public Spending on Education (G8 Countries)",
    x = "Public Spending (% of GDP)",
    y = "Youth Unemployment (%)",
    color = "Education",
    shape = "Sex",
    caption = "Source: ILO/World Bank"
  ) +
  theme_minimal()

print(scatter_plot_g8)
```
#### b) By sex 
```{r}
linear_model_g8_sex <- lm(`Youth unemployment (%)` ~ `Public Spending (% of GDP)` + sex, data = dep_chomage_g8_filtered)

summary(linear_model_g8_sex)

scatter_plot_g8_sex <- ggplot(dep_chomage_g8_filtered, aes(x = `Public Spending (% of GDP)`, y = `Youth unemployment (%)`,color = sex)) +
  geom_point() +
  facet_wrap(~sex)+
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(
    title = "Figure 19 : Youth Unemployment vs Public Spending by sex (G8 Countries)",
    x = "Public Spending (% of GDP)",
    y = "Youth Unemployment (%)",
    color = "Sex",
    caption = "Source: ILO/World Bank"
  ) +
  theme_minimal()

print(scatter_plot_g8_sex)
```

#### c) Education and sex 

#### Regression Model for G8 countries 

The simple linear regression model can be represented as:

\[
 \ Y_{t} = ß_{0} + ß_{1}X_{t} + ß{2}*Education + ß{3}*sex   \varepsilon_{t} \varepsilon_{t}
\]

where:
- \(Y_{t}\): Youth unemployment rate at time \(t\).
- \(\beta_{0}\): Intercept term.
- \(\beta_{1}\): Coefficient for the variable \(X_{t}\).
- \(X_{t}\): Public spending variable at time \(t\).
- \(\epsilon_{t}\): Error term.


```{r}
linear_model_g8_sex_education <- lm(`Youth unemployment (%)` ~ `Public Spending (% of GDP)` + Education + sex, data = dep_chomage_g8_filtered)

summary(linear_model_g8_sex)

scatter_plot_g8 <- ggplot(dep_chomage_g8_filtered, aes(x = `Public Spending (% of GDP)`, y = `Youth unemployment (%)`, color = Education, shape = sex)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color="red") +
  labs(
    title = "Figure 19 : Youth Unemployment vs Public Spending by sex and level of education (G8 Countries)",
    x = "Public Spending (% of GDP)",
    y = "Youth Unemployment (%)",
    color = "Education",
    shape = "Sex",
    caption = "Source: ILO/World Bank"
  ) +
  theme_minimal()

print(scatter_plot_g8)
``` 

# Conclusion 



# Annexe

##Github repository The GitHub project can be found on: *https://github.com/GhinaMezher/project-BM*.git 2 collaborators can be found on this public project.

## Links and sources

We used the data base from the World Bank and International Labor Organization that identifies the public spendings in education: *https://data.worldbank.org/indicator/SE.XPD.TOTL.GD.ZS?end=2022&start=1970&view=chart* And for the data on youth unemployment rates depending on various caracteristics we referred to the International Labor Organisation (ILO) database: *https://ilostat.ilo.org/fr/topics/youth/#* & *https://www.ilo.org/shinyapps/bulkexplorer27/?lang=fr&id=POP_3WAP_SEX_AGE_EDU_NB_A*


